---
title: "Data Cleaner"
author: "Daniel K Baissa"
date: "1/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## Galager's Dataset

Election indices
The figures below represent the values of three indices:
(i) the least squares index (LSq), which measures disproportionality between the
vote distribution and the seat distribution;
(ii) the effective number of parties at the electoral level (Eff Nv, also termed
ENEP);
(iii) the effective number of parties at the parliamentary or legislative level (Eff Ns,
also termed ENPP).

The N of seats refers to the number given in the sources used and is the number on
which the calculations were based (though see point (i) below).
These indices were originally outlined in
Markku Laakso and Rein Taagepera, ‘ “Effective” number of parties: a measure with
application to west Europe’, Comparative Political Studies 12:1 (1979), pp. 3–27
(effective number of parties), and
Michael Gallagher, ‘Proportionality, disproportionality and electoral systems’,
Electoral Studies 10:1 (1991), pp. 33–51 (least squares index).
Details of these indices, and of how they have been calculated, can be found in
Appendix B of:
Michael Gallagher and Paul Mitchell (eds), The Politics of Electoral Systems
paperback edition (Oxford and New York: Oxford University Press, 2008).
For further details on this book, see :
https://global.oup.com/academic/product/the-politics-of-electoral-systems-
9780199238675?cc=ie&lang=en&
The main historical sources for the election results from which these indices are
calculated are:
Thomas T. Mackie and Richard Rose, The International Almanac of Electoral
History, 3rd ed (Basingstoke: Macmillan, 1991);
Annual Data Section in European Journal of Political Research since 1990;
Richard Rose and Neil Munro, Elections and Parties in New European
Democracies (Washington: CQ Press, 2003);
Dieter Nohlen, Michael Krennerich and Bernhard Thibaut (eds), Elections in
Africa: a data handbook (Oxford: Oxford University Press, 1999);
Dieter Nohlen, Florian Grotz and Christof Hartmann (eds), Elections in Asia and
the Pacific: a data handbook, 2 vols (Oxford: Oxford University Press, 2001);
Dieter Nohlen (ed.), Elections in the Americas: a data handbook, 2 vols (Oxford:
Oxford University Press, 2005).
More recent election results, and indeed some earlier ones, are drawn from a range of
internet sites, where possible official ones such as an election commission or electoral
tribunal. A list of some of these is given in Appendix E of Gallagher and Mitchell
(eds), The Politics of Electoral Systems. Other useful sites, not listed there, include:
www.caribbeanelections.com (very informative on all aspects of Caribbean
politics, at least up to 2019)
africacheck.org/
africanelections.tripod.com (active until around 2012)
Election indices 2
While any user of printed and internet sources must appreciate the time that has been
spent in compiling these and the care taken to ensure accuracy, the main problems
associated with sources (and, implicitly, the main appeals to those who compile
election results) are:
(i) bunching of ‘Others’, i.e. small parties and independents not listed separately.
In the calculation of indices, the greater the amount of disaggregation in the data, the
better. Ideally, every party winning more than 0.1 per cent of the national vote,
certainly 0.5 per cent, should be listed separately. The votes of Independents are very
rarely disaggregated, so when they are a significant force (Japan, South Korea,
Maldives, Ukraine and sometimes other countries too), there are problems in trying to
compute indices given that each independent candidate must be treated as a separate
‘party’. The approach taken here in such cases has been that outlined in the
IndicesCalc file at
https://www.tcd.ie/Political_Science/people/michael_gallagher/ElSystems/Docts/IndicesCalc.pdf
Fortunately, many countries now provide ‘perfect’ data, i.e. complete disaggregation
down to the level of each individual independent candidate.
(ii) occasionally, the problem is the opposite of (i), namely that results exist only in
disaggregated form that no-one has yet taken the trouble to aggregate (or to make the
figures generally available if they have been aggregated), as with Argentina or the
constituency-level votes at a number of elections in Lithuania. (In Argentina, the
different parties (or party names) and alliances in different provinces can make it very
difficult to identify reliable national-level vote figures, quite apart from the problem
of incomplete data.) This is a problem particularly in mixed parallel systems (also
known as mixed-member majoritarian systems, or MMM systems) in which
constituency seats are allocated on the basis of constituency votes and list votes on
the basis of list votes, with no linkage between the two. While national-level figures
for the list votes are almost always available, that is not always the case for the
constituency votes, or, if they are available, they have not been aggregated and the
researcher would have to embark on the sizeable task of putting together overall
results from the results in dozens of individual constituencies. This is a problem for
some elections in, for example, Georgia, South Korea, Lithuania and Ukraine.
(iii) occasional logical inconsistency, i.e. the number of votes or seats for the listed
parties does not add to the stated total, or a party with no votes is stated to have won
seats. A list of corrections that need to be made to the Mackie and Rose figures can be
found in Arend Lijphart, Electoral Systems and Party Systems: A Study of Twenty-
Seven Democracies, 1945–1990 (Oxford and New York: Oxford University Press,
1994), pp. 163–77. Generally, the approach adopted here has been the same as
Lijphart’s, i.e. when the reported number of total valid votes (or seats) does not equal
the sum of the reported votes (or seats) for individual parties, the number used as the
basis for calculations has been the sum of the parties’ votes (or seats)
(iv) absence of official results in Roman script. While, understandably, the main
audience for national election results consists of citizens of that country, the absence
of results in Roman script for some well-resourced countries such as Japan, the
Republic of Korea and Taiwan makes the reconstitution of results from these
Election indices 3
countries quite time-consuming and sometimes impossible for those who are able to
read only Roman script.
An issue that arises occasionally under mixed systems (those where some candidates
are elected from small constituencies and others from a national or regional list) is
that some constituency seats are won by independent candidates or small parties that
do not run in the list element of the election. (Examples include Egypt, Morocco,
Nepal, Palestine.) When only list vote totals are available, or when list votes alone are
the basis for allocating seats to parties, this creates a situation where it appears,
comparing national votes with seats, that somehow a party that won no votes
nonetheless won a seat. Anomalous as this is, it seems preferable to simply ignoring
these candidates / parties and basing the calculations on the other seats, which would
distort every other party’s percentage of the seats. In cases where the number of seats
won by parties or candidates that did not run in the list element is large (for example,
Ukraine’s election of 2012, where 49 of the 225 constituency seats were won by
small parties or independents that won no list votes and for which there is no data on
constituency votes), the election is not included in the dataset.
The countries included here include the standard set that tend to feature in
comparative politics analyses and, relatedly of course, are covered in the historical
sources listed above. As many other countries as possible are also included for
particular elections, dependent on the availability of reliable and adequately
disaggregated results. The elections included are confined to those that could
plausibly be deemed occasions of choice for the voting population, which is not to
imply that every single election included here can be regarded as meeting the highest
democratic standards.
A complication that affects the calculation of Eff Ns in a few cases is that groups of
parties may contest an election as a coalition in order to reap the benefits of size and
then disaggregate after the election. For example, in Greece’s 1958 election only five
groupings contested the election and have recorded vote totals, but because two of
these groupings were umbrellas, seat totals are recorded for nine different parties
(Mackie and Rose, International Almanac, pp. 198–200). In the figures below, Eff Ns
is based on seat totals for the same groups as are used for the calculation of Eff Nv,
disregarding any post-election splitting of the groupings that contested the election
together. In a number of elections in some other countries such as Italy, Malaysia and
San Marino, figures are available both for individual parties and for the broader
alliances into which some of them group; in most cases, the figures in the tables are
based on treating individual parties as the units, with figures using alliances given in
the Notes.
Conditions of use: there aren’t any, but citations are always appreciated ...
Suggested citation format:
Gallagher, Michael, 2021. Election indices dataset at
http://www.tcd.ie/Political_Science/people/michael_gallagher/ElSystems/index.php,
accessed [date].

```{r}
library(readxl)

sheet_names <- excel_sheets("Original_data/IndicesWd_ElectionIndices.xlsx")           # Get sheet names
sheet_names 

```
This takes the excel sheets and combines them into a tibble.

```{r}
list_all <- lapply(sheet_names, function(x) {as.data.frame(read_excel("Original_data/IndicesWd_ElectionIndices.xlsx", sheet = x)) } )
```

